Thesis Proposal

### 2019-09-20 ThesisProposal.Rmd
Same but with spell check

Currently, “there is no clear-cut method for determining when the MICE algorithm has converged” (@buur18, \textsection 6.5). Instead, users have to rely on visual inspection of the algorithm's iterations. Convergence is said to be reached when parameters (like means of completed variables or regression coefficients) are stable across iterations (@whit11). 
<!-- "As MICE is an iterative procedure, it is important that convergence is achieved. This may be checked by computing, at each cycle, the means of imputed values and/or the values of regression coefficients, and seeing if they are stable" (@whit11, p. 394). -->
And additionally, the variation between imputations should be no larger than the variation across iterations of each individual chain (@buur18). 
<!-- “Convergence is diagnosed when the variance between different sequences is no larger than the variance within each individual sequence” (@buur18, \textsection 6.5).  -->
*Conceptually, this visual evaluation procedure is similar to the 'Gelman-Rubin statistic' (@li14). Practically, however, Gelman and Rubin's convergence diagnostic $\widehat{R}$ ('potential scale reduction'; 1992) cannot be directly applied to assess convergence of MICE algorithms because the measure is famously prone to produce false positives when applied to MI data.*  
<!-- "A common diagnostic tool is to plot one or more parameters against the iteration number and assess convergence by how different the variance between different sequences is relative to the variance within each individual sequence, similar to the Gelman-Rubin statistic (Gelman and Rubin, 1992) used in Markov chain Monte Carlo (MCMC) diagnostics" (@li14, \textsection 4.3). -->
*That is, the overdispersed initial state of the MICE algorithm does not conform to $\widehat{R}$'s assumption of equal starting points (source!).*

Notwithstandingly, it seems like @su11 did implement $\widehat{R}$ as convergence criterion into their R package 'mi', including the conventional cut-off of 1.1.
<!-- "R.hat: The value of the $\hat{R}$ statistic used as a convergence criterion. The default is 1.1 (Gelman and Rubin 1992; Gelman, Carlin, Stern, and Rubin 2004)" (@su11 p. 4). -->
<!-- "Our mi offers two ways to check the convergence of the multiple imputation procedure. By default, mi() monitors the mixing of each variable by the variance of its mean and standard deviation within and between different chains of the imputation. If the $\hat{R}$ statistic is smaller than 1.1, (i.e., the difference of the within and between variance is trivial), the imputation is considered converged (Gelman, Carlin, Stern, and Rubin 2004). Additionally, by specifying mi(data, check.coef.convergence = TRUE, ...), users can check the convergence of the parameters of the conditional models" (@su11, p. 13). -->
It is worth investigating the validity of the convergence statistic in the context of MICE, and the broader framework of fully conditional specification (FSC) algorithms.

@@

The research report that is to be handed in December 13th 2019 will consist of an investigation into algorithmic convergence of MICE. Ideally, this will result in a single indicator to flag non-convergence (potentially based on the Gelman-Rubin statistic, autocorrelation, or MC error).

The second stage of the research project is of more practical nature: programming ShinyMICE. The application will at least consist of the following: 1) one or more measures to assess algorithmic convergence; 2) data visualizations (e.g., scatterplots, densities, and cross-tabulations of the data pre- and post-imputation); and 3) statistical evaluation of relations between variables pre- versus post-imputation (i.e., chi square tests or t-tests). This part of the thesis also entails research into user interface (UI) design, local versus cloud-based data storage, and optimizing the application's efficiency.

@@

>>>>

Currently, “there is no clear-cut method for determining when the MICE algorithm has converged” (@buur18, \textsection 6.5). Instead, users have to rely on visual inspection of the algorithm's iterations. Convergence is said to be reached when parameters (like means of completed variables or regression coefficients) are stable across iterations (@whit11). 
<!-- "As MICE is an iterative procedure, it is important that convergence is achieved. This may be checked by computing, at each cycle, the means of imputed values and/or the values of regression coefficients, and seeing if they are stable" (@whit11, p. 394). -->
And additionally, the variation between imputations should be no larger than the variation across iterations of each individual chain (@buur18). 
<!-- “Convergence is diagnosed when the variance between different sequences is no larger than the variance within each individual sequence” (@buur18, \textsection 6.5).  -->
*Conceptually, this visual evaluation procedure is similar to the 'Gelman-Rubin statistic' (@li14). Practically, however, Gelman and Rubin's convergence diagnostic $\widehat{R}$ ('potential scale reduction'; 1992) cannot be directly applied to assess convergence of MICE algorithms because the measure is famously prone to produce false positives when applied to MI data.*  
<!-- "A common diagnostic tool is to plot one or more parameters against the iteration number and assess convergence by how different the variance between different sequences is relative to the variance within each individual sequence, similar to the Gelman-Rubin statistic (Gelman and Rubin, 1992) used in Markov chain Monte Carlo (MCMC) diagnostics" (@li14, \textsection 4.3). -->
*That is, the over-dispersed initial state of the MICE algorithm does not conform to $\widehat{R}$'s assumption of equal starting points (source?).*

Notwithstanding, it seems like @su11 did implement $\widehat{R}$ as convergence criterion into their R package 'mi', including the conventional cut-off of 1.1.
<!-- "R.hat: The value of the $\hat{R}$ statistic used as a convergence criterion. The default is 1.1 (Gelman and Rubin 1992; Gelman, Carlin, Stern, and Rubin 2004)" (@su11 p. 4). -->
<!-- "Our mi offers two ways to check the convergence of the multiple imputation procedure. By default, mi() monitors the mixing of each variable by the variance of its mean and standard deviation within and between different chains of the imputation. If the $\hat{R}$ statistic is smaller than 1.1, (i.e., the difference of the within and between variance is trivial), the imputation is considered converged (Gelman, Carlin, Stern, and Rubin 2004). Additionally, by specifying mi(data, check.coef.convergence = TRUE, ...), users can check the convergence of the parameters of the conditional models" (@su11, p. 13). -->
It is worth investigating the validity of the convergence statistic in the context of MICE, and the broader framework of fully conditional specification (FSC) algorithms.

@@

The research report that is to be handed in December 13th 2019 will consist of an investigation into algorithmic convergence of MICE. Ideally, this will result in a single indicator to flag non-convergence (potentially based on the Gelman-Rubin statistic, auto-correlation, or MC error).

The second stage of the research project is of more practical nature: programming ShinyMICE. The application will at least consist of the following: 1) one or more measures to assess algorithmic convergence; 2) data visualizations (e.g., scatter-plots, densities, and cross-tabulations of the data pre- and post-imputation); and 3) statistical evaluation of relations between variables pre- versus post-imputation (i.e., chi square tests or t-tests). This part of the thesis also entails research into user interface (UI) design, local versus cloud-based data storage, and optimizing the application's efficiency.

@@

"Why can the number of iterations in MICE be so low? First of all, realize that the imputed data Ymis form the only memory in the the MICE algorithm. Chapter 3 explained that imputed data can have a considerable amount of random noise, depending on the strength of the relations between the variables. Applications of MICE with lowly correlated data therefore inject a lot of noise into the system. Hence, the auto-correlation over t will be low, and convergence will be rapid, and in fact immediate if all variables are independent. Thus, the incorporation of noise into the imputed data has pleasant side-effect of speeding up convergence" (@buur18, par. 4.5).
